---
title: "PhyloTrees"
author: "Tessa Lobnow"
date: "11/2/2021"
output: html_document
---

``` {r load data, echo = F, message = F, warning = F}
#install.packages("adegenet")
#install.packages("phangorn")

library(adegenet)
library(phangorn)
library(stats)
library(ade4)
library(ape)


```



# Tree Construction Methods

## Distance-Based

- find genetic distance between pairs of taxa + cluster species by distances
- Pairs w/ shorter distances cluster more closely than pairs w/ larger distances
- can be rooted/unrooted

Advantages:
+ fast
+ flexible

Disadvantages:
- tree can change based on algorithm used
- no testing for best model possible
- tree can be inaccurate





## Maximum Parsimony

- aim to minimize # of character-state changes
- simplest possible tree w/ least branches
- Composition:
  1) initiates algorithm for using a tree
  2) makes small changes by simplifying relationships until we have the most parsimonious tree (no further simplifiable)
  
  
  
  

## Likelihood-based

- e.g. Bayesian methods
- Max. likelihood maximizes the chances to collecting the data set in question
  - given that the calculated tree are true (along with parameters describing evolutionary rules)

- Mathematically: P(D | M)
  - P ... Probability
  - D ... Observed Data
  - M ... Model
  
- e.g.: use an observed set of coin flips to say whether this outcome is more likely under the model that the coin is fair or biased

- favors minimizing the # of mutations at "internal nodes" (divergence points) inside a tree
- based on the assumption that the fewest # of possible substitutions to get from seq.version A to seq.version B is most likely what happened (not always displays reality)


``` {r load FASTA sequences, echo = F, message = F, warning = F}

dna <- fasta2DNAbin(file = "/Users/FinnLo/Documents/Programming/R/HZ_SC_and_Raw_Data/Cryptosporidium-BSc/FASTA_sequences/MSC6-7_seqs.fasta")

dna


D <- dist.dna(dna, model = "TN93")
length(D) #number of pairwise distances, computed as n(n-1)/2

temp <- as.data.frame(as.matrix(D))
temp <- temp[1:9, 1:9]
table.paint(df = temp,  
  temp, cleg=0, clabel.row=0.5, clabel.col=0.5) 
# darker shades of gray mean a larger distance 
# you can also make cool color plots but they're much more complicated because they use the image() function
```






